Demo arXiv template

### 2022-10-02 paper/paper.qmd
build up more content

Numerous indices have been proposed by researchers and practitioners to monitor natural hazard, for example, @alahacoon_comprehensive_2022 reviews 111 drought indices derived from traditional and remote sensing data; [review of index construction in other area, climate indices, many other reviews in drought].

@@

# Data pipeline

Constructing a pipeline that divides a complex procedure into steps that can be concatenated has been adopted widely in the computational statistics community. 
One particular example is a machine learning pipeline, tidymodel [@tidymodel]. 
[more details on pipeline]

**Statistical pipeline in interactive graphics**

The data pipeline in interactive graphics is a set of steps that transform the raw data to the plots displayed on the screen. 
The initial pipeline proposed by @buja_elements_1988 involves the following steps: non-linear transformation, variables standardization, randomization, projection engine, and viewporting.
Another example in the early work of pipeline by @sutherland_orca_2000 describes a three-step pipeline: variable standardization, dimension reduction, and scaling data into the viewing window. 
This pipeline also includes the transformation on spatial and temporal variables, i.e. computing time lag on temporal variables.
@wickham_plumbing_2009 argues that whether made explicit or not, pipeline has to be presented in every graphics program and breaking down graphic rendering into steps is also beneficial for understand the implementation and compare between different graphic systems.

**Statistical pipeline in other domains**

TODO: read papers on other pipeline in R

# A pipeline for natural hazard indices

@@

>>>>

@@

# Data pipeline in R

Constructing a pipeline that divides a complex procedure into steps that can be concatenated has been adopted widely in the R community.

The data pipeline in interactive graphics is a set of steps that transform the raw data to the plots displayed on the screen. 
The initial pipeline proposed by @buja_elements_1988 involves the following steps: non-linear transformation, variables standardization, randomization, projection engine, and viewporting.
The initial pipeline proposed by @buja_elements_1988 involves the following steps: non-linear transformation, variables standardization, randomization, projection engine, and viewporting. 
Another example in the early work of pipeline by @sutherland_orca_2000 describes a three-step pipeline: variable standardization, dimension reduction, and scaling data into the viewing window. 
This pipeline also includes the transformation on spatial and temporal variables, i.e. computing time lag on temporal variables.
This pipeline also includes the transformation on spatial and temporal variables, i.e. computing time lag on temporal variables. 
@wickham_plumbing_2009 argues that whether made explicit or not, pipeline has to be presented in every graphics program and breaking down graphic rendering into steps is also beneficial for understand the implementation and compare between different graphic systems.

The data pipeline concept is further enhanced by the pipe operator (`%>%`) in R where a set of operations, or steps, can be chained together to form a set of instructions.

A more recent data pipeline is tidymodels [@tidymodels], a set of packages for machine learning models following the tidyverse principles [@tidyverse]. \[expand on tidymodels\]

# A pipeline for building natural hazard indices

The construction of natural hazard indices also follows a set of steps, which is usually illustrated using a flowchart in the paper. 
However, every researcher follows a certain design philosophy and steps taken in the index constructed by different researchers are not aligned. 
This discourages experiment with multiple indices. Initiate a new workflow when computing a new index.

## Raw data

The data used to construct the natural hazard index usually have three dimensions, one for location, one for time, and one for multivariate. 
Mathematically, it can be written as $X_{j, s, t}$, where $j = 1, 2, \cdots, J$ for variable, $s = 1, 2, \cdots, S$ for location, and $t = 1, 2, \cdots, T$ for time.

The location $s$ can refer to vector points or areas characterised by longitude-latitude coordinates, or raster cells obtained from satellite images.

The time dimension $t$ can be daily, weekly, biweekly (14-16 days), monthly, or even quarterly

Variables

This multidimensional array structure is commonly used in geospatial analysis

Given the variety of data sources at different spatial resolution and temporal granularity, the raw data may first come in multiple pieces. 
Sometimes, even a considerable amount of work is needed to align the spatial and temporal extent of multivariate data.

A notation for different variables have different spatial and temporal granularity $X_{j_1, s_1, t_1}$???

## Spatial aggregation

## Variable transformation

Take one or multiple variables to create a new variable, linearly? 
sometimes called feature extraction in the machine learning community 
With drought indices, the extraction of meaningful variables from the original data is usually supported by the water balance model, for example, in SPEI, the step that create $d$ out of precipitation and potential evapotranspiration (PET) has theoretical backup from [see paper.]

## Temporal processing

## Dimension reduction

## Normalising

only for distribution fit, convert from an fitted distribution to normal

## Benchmarking

## Simplification

@@

# Extending the pipeline

### 2022-10-02 paper/paper.tex
build up more content

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi

@@

\hypertarget{data-pipeline}{%
\section{Data pipeline}\label{data-pipeline}}

Constructing a pipeline that divides a complex procedure into steps that
can be concatenated has been adopted widely in the computational
statistics community. One particular example is a machine learning
pipeline, tidymodel (Kuhn and Wickham 2020). {[}more details on
pipeline{]}

\textbf{Statistical pipeline in interactive graphics}

@@

\textbf{Statistical pipeline in other domains}

TODO: read papers on other pipeline in R

\hypertarget{a-pipeline-for-natural-hazard-indices}{%
\section{A pipeline for natural hazard
indices}\label{a-pipeline-for-natural-hazard-indices}}

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-tidymodel}{}}%
Kuhn, Max, and Hadley Wickham. 2020. \emph{Tidymodels: A Collection of
Packages for Modeling and Machine Learning Using Tidyverse Principles.}
\url{https://www.tidymodels.org}.

@@

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

@@

\hypertarget{data-pipeline-in-r}{%
\section{Data pipeline in R}\label{data-pipeline-in-r}}

Constructing a pipeline that divides a complex procedure into steps that
can be concatenated has been adopted widely in the R community.

@@

The data pipeline concept is further enhanced by the pipe operator
(\texttt{\%\textgreater{}\%}) in R where a set of operations, or steps,
can be chained together to form a set of instructions.

A more recent data pipeline is tidymodels (Kuhn and Wickham 2020), a set
of packages for machine learning models following the tidyverse
principles (Wickham et al. 2019). {[}expand on tidymodels{]}

\hypertarget{a-pipeline-for-building-natural-hazard-indices}{%
\section{A pipeline for building natural hazard
indices}\label{a-pipeline-for-building-natural-hazard-indices}}

The construction of natural hazard indices also follows a set of steps,
which is usually illustrated using a flowchart in the paper. However,
every researcher follows a certain design philosophy and steps taken in
the index constructed by different researchers are not aligned. This
discourages experiment with multiple indices. Initiate a new workflow
when computing a new index.

\hypertarget{raw-data}{%
\subsection{Raw data}\label{raw-data}}

The data used to construct the natural hazard index usually have three
dimensions, one for location, one for time, and one for multivariate.
Mathematically, it can be written as \(X_{j, s, t}\), where
\(j = 1, 2, \cdots, J\) for variable, \(s = 1, 2, \cdots, S\) for
location, and \(t = 1, 2, \cdots, T\) for time.

The location \(s\) can refer to vector points or areas characterised by
longitude-latitude coordinates, or raster cells obtained from satellite
images.

The time dimension \(t\) can be daily, weekly, biweekly (14-16 days),
monthly, or even quarterly

Variables

This multidimensional array structure is commonly used in geospatial
analysis

Given the variety of data sources at different spatial resolution and
temporal granularity, the raw data may first come in multiple pieces.
Sometimes, even a considerable amount of work is needed to align the
spatial and temporal extent of multivariate data.

A notation for different variables have different spatial and temporal
granularity \(X_{j_1, s_1, t_1}\)???

\hypertarget{spatial-aggregation}{%
\subsection{Spatial aggregation}\label{spatial-aggregation}}

\hypertarget{variable-transformation}{%
\subsection{Variable transformation}\label{variable-transformation}}

\hypertarget{temporal-processing}{%
\subsection{Temporal processing}\label{temporal-processing}}

\hypertarget{dimension-reduction}{%
\subsection{Dimension reduction}\label{dimension-reduction}}

\hypertarget{normalising}{%
\subsection{Normalising}\label{normalising}}

\hypertarget{benchmarking}{%
\subsection{Benchmarking}\label{benchmarking}}

\hypertarget{simplification}{%
\subsection{Simplification}\label{simplification}}

@@

\hypertarget{extending-the-pipeline}{%
\section{Extending the pipeline}\label{extending-the-pipeline}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-tidymodels}{}}%
Kuhn, Max, and Hadley Wickham. 2020. \emph{Tidymodels: A Collection of
Packages for Modeling and Machine Learning Using Tidyverse Principles.}
\url{https://www.tidymodels.org}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-tidyverse}{}}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.
